\begin{comment}
3 Calculating datastructures using Ornaments

In this part we return to the matter numerical representations. With 2.3 in mind, we can rephrase part our original question to ask

> Can numerical representations be described as ornaments on their number systems?

Let us look at a numerical representation presented as ornament in action.

\section{Numerical representations as ornaments}\label{sec:ornaments}
Reflecting on this derivation for \bN{}, we could perform the same computation for \bL{} to get Braun trees. However, we note that these computations proceed with roughly the same pattern: each constructor of the numeral system gets assigned a value, and is amended with a field holding a number of elements and subnodes using this value as a ``weight''. This kind of ``modifying constructors'' is formalized by ornamentation \cite{progorn}, which lets us formulate what it means for two types to have a ``similar'' recursive structure. This is achieved by interpreting (indexed inductive) datatypes from descriptions, between which an ornament is seen as a certificate of similarity, describing which fields or indices need to be introduced or dropped to go from one description to the other. \textit{Ornamental descriptions}, which act as one-sided ornaments, let us describe new datatypes by recording the modifications to an existing description.
\todo[inline]{Put some minimal definitions here.}

Looking back at \AgdaDatatype{Vec}, ornaments let us show that express that \AgdaDatatype{Vec} can be formed by introducing indices and adding a fields holding an elements to \bN{}.
However, deriving \AgdaDatatype{List} from \bN{} generalizes to \bL{} with less notational overhead, so we tackle that case first. We use the following description of \bN{}
\ExecuteMetaData[Tex/NumRepOrn]{NatD}
Here, \AgdaInductiveConstructor{σ} adds a field to the description, upon which the rest of the description can vary, and \AgdaInductiveConstructor{ṿ} lists the recursive fields and their indices (which can only be \AgdaInductiveConstructor{tt}).
We can now write down the ornament which adds fields to the \AgdaFunction{suc} constructor
\ExecuteMetaData[Tex/NumRepOrn]{ListO}
Here, the \AgdaInductiveConstructor{σ} and \AgdaInductiveConstructor{ṿ} are forced to match those of \AgdaDatatype{NatD},
but the \AgdaInductiveConstructor{Δ} adds a new field. Using the least fixpoint and description extraction, we can then define \AgdaDatatype{List} from this ornamental description. Note that we cannot hope to give an unindexed ornament from \bL{}
\ExecuteMetaData[Tex/NumRepOrn]{LeibnizD}
into trees, since trees have a very different recursive structure! Thus, we must keep track at what level we are in the tree so that we can ask for adequately many elements:
\ExecuteMetaData[Tex/NumRepOrn]{TreeO}
We use the \AgdaFunction{power} combinator to ensure that the digit at position $n$, which has weight $2^n$ in the interpretation of a binary number, also holds its value times $2^n$ elements. This makes sure that the number of elements in the tree shaped after a given binary number also is the value of that  binary number.
\end{comment}


\section{Generic numerical representations}\label{sec:trieo}
In this section, we will demonstrate how we can use ornamental descriptions to generically compute numerical representations. 

The reasoning here proceeds differently from that in the calculation of \AD{Vec} from \bN{}. Indeed, we directly construct datatypes, 
%and only prove it is the correct type after,
rather than deriving them step-by-step using isomorphism reasoning. Nevertheless, the choices of fields depending on the analysis of a number system follow the same strategy. We will first present the unindexed numerical representations, explaining which fields it adds and why, by cases on the number system. Then, we will show the indexed numerical representations as an ornament on top of the unindexed variant, and how the indices built up incrementally as we descend over the structure of the number system.

Recall the ``natural numbers''-information \AF{Number}, which gets its semantics from the conversion to \bN{}:
\ExecuteMetaData[Ornament/Numerical]{toN-type}
which is defined by generalizing over the inner information bundle and folding using
\ExecuteMetaData[Ornament/Numerical]{toN-con}
The choice of interpretation restricts the numbers to the class of numbers which are evaluated as linear combinations of ``digits''\footnote{An arbitrary \AF{Number} system is not necessarily isomorphic to \bN{}, as the system can still be incomplete (i.e., it cannot express some numbers) or redundant (it has multiple representations of some numbers).}. This class certainly does not include all interesting number systems, but does include many systems that have associated arrays\footnote{Notably, arbitrary polynomials also have numerical representations, interpreting multiplication as precomposition.}. 

We let this interpretation into \bN{} guide the construction of the associated numerical representation. In each case, we follow the computation in \AF{value} by inserting vectors of sizes corresponding to the weights of the number system:
\ExecuteMetaData[Ornament/Numerical]{trieifyOD}
In\todo{Explain better} the case of a leaf \AIC{𝟙} of weight \AV{k}, we insert a vector of size \AV{k}. Similarly, in a field \AIC{σ}, where the weight is determined by a value \AV{s} of \AV{S}, we insert a vector of the weight corresponding to the value of \AV{s}. Note that the actual value/number of elements a leaf or field contributes depends on the preceding multipliers of recursive fields: a recursive field of a number can have a weight \AV{k}, so we multiply the number of elements in a recursive sequence by wrapping the parameter in a vector of size \AV{k}. By roughly the same reasoning we pass the trieification of a subdescription \AV{R} the parameter wrapped in a vector, which we compose into the current numerical description by using the ornament \AIC{∙δ}. Since \AV{R} can have a different \AD{Info}, we generalized the whole construction over \AV{ϕ}\ \AV{:} \AD{InfoF}\ \AV{If}\ \AF{Number}.

As an example, let us define \AF{PhalanxD} as a number system and walk through the computation of its \AF{trieifyOD}. We define
\ExecuteMetaData[Ornament/Numerical]{PhalanxND}
Now, we see that applying \AF{trieifyOD} sends leaves with a value of \AV{k} to \AD{Vec}\ \AV{A}\ \AV{k}, so applying it to \AF{DigitND} yields
\ExecuteMetaData[Ornament/Numerical]{DigitOD-2}
which is equivalent to the \AF{DigitOD} from before, up to expanding a vector of \AV{k} elements into \AV{k} fields. The same happens for the first two constructors of \AF{PhalanxND}, replacing them with an empty vector and a one-element vector respectively. The \AF{ThreeND} in the last constructor gets trieified to \AF{DigitOD′} and composed by \AF{O∙δ+}, and the recursive field gets replaced by a recursive field nesting over vectors of length. Again, this is equivalent to \AF{FingerOD}, up to wrapping values in length one vectors, replacing \AD{Pair} with a two-element vector, and inserting empty vectors.

The indexed numerical representations can be constructed from the unindexed numerical representations\todo{Explain better} by adding fields to track the indices where necessary, incrementally combining these into the indices at the leaves. In contrast to the computation of \AD{Vec} in \autoref{sec:desc-numrep}, where we gave the definition of the datatype by cases on the index, we will have to track the indices as we descend over structure of the number system by generalizing over an index-computing algebra: 
\ExecuteMetaData[Ornament/Numerical]{itrieifyOD}
Explain\todo{This concludes a bunch of things, including this thesis.}

